### 2015-09-12 vmcai-2016-tex/abstract.tex
Revision and refactor
edits to evaluation and conclusion

A recent trend in symbolic execution of object-oriented programs is the modeling of references as sets of guarded values, enabling multiple heap shapes to be represented in a single state. A fundamental problem with using these guarded value sets is the creation of test inputs for programs accepting symbolic reference input parameters. Although several solutions have been proposed, none have been proven to be sound and complete with respect to the properties provable by generalized symbolic execution (GSE). This work presents a method for initializing reference inputs in a black-box symbolic input heap that exactly preserves GSE semantics. A correctness proof for the initialization scheme is provided, as well as the results of an empirical evaluation of a proof-of-concept implementation. The initialization technique can be used to ensure that guarded value set based symbolic execution engines operate in a provably correct manner with regards to symbolic references.

>>>>

A recent trend in the analysis of object-oriented programs is the
modeling of references as sets of guarded values, enabling multiple
heap shapes to be represented in a single state. A fundamental problem
with using these guarded value sets is the creation of test inputs for
programs accepting symbolic reference input parameters. Although
several solutions have been proposed, none have been proven to be
sound and complete with respect to the properties provable by
generalized symbolic execution (GSE). This work presents a method for
initializing reference inputs in a symbolic input heap using guarded
value sets that exactly preserves GSE semantics. A correctness proof
for the initialization scheme is provided, as well as the results of
an empirical evaluation of a proof-of-concept implementation. The
initialization technique can be used to ensure that guarded value set
based symbolic execution engines operate in a provably correct manner
with regards to symbolic references.

### 2015-09-12 vmcai-2016-tex/evaluation.tex
edits to evaluation and conclusion

\section{Evaluation}
\label{sec:evaluation}
%\begin{figure*}[t]
\begin{table*} [t]
  \centering
  \scalebox{0.88}{\begin{tabular}{| c | c | r | r | r | r | r | r | r | r | r |}
  \hline
   \multirow{2}{*}{Method }&\multirow{2}{*}{ $k$ }
   &\multicolumn{3}{c|}{Time (seconds)} &\multicolumn{3}{c|}{States} &\multicolumn{3}{c|}{ Paths }\\
								&	&\gsetxt{} & L\#		&SL		&\gsetxt{}	& L\# & SL&\gsetxt{}	& L\# 	& SL\\
   \hline
    \multirow{3}{*}{LinkedList }			&3	& 0.91	& 1.21	& 0.69	& 2465	& 2844	& 99		&1656	& 1269	& 25\\
   		 						& 4	& 2.92	& 3.35	& 0.91	& 25774	& 29977	& 155	&17485	& 13550	& 39\\
   								& 5	& 20.78	& 19.47	& 1.59	& 341164	& 400296	& 223	&232743	& 181849	& 56\\
								& 6	& 280.56	& 299.19	& 2.36	&5447980	&6437201	& 303	&3731094	&2933027	& 76\\
								& 7	& -		& -		& 5.07	& -		&-		& 395	&-		&-		& 99\\
								& 8	& -		& -		& 17.49	& -		&-		& 499	&-		&-		& 125\\
								& 9	& -		& -		& 63.96	& -		&-		& 615	&-		&-		& 154\\
								& 10	& -		& -		& 206.93	& -		&-		& 743	&-		&-		& 186\\
    \hline
    \multirow{3}{*}{BinarySearchTree }	& 1	& 0.26	& 0.28	& 0.36	& 19		& 23		& 29		& 6		& 6		& 6\\
   		 						& 2	& 0.83	& 1.28	& 0.93	& 143	& 143	& 145	& 43		& 42		& 33\\
   								& 3	& 20.63	& 25.55	& 4.03	& 1953	& 1703	& 1485	& 515	& 515	& 328\\
								& 4	& -		& -		& 410.89	& -		&-		& 73635	&-		&-		& 15563\\
    \hline
      \multirow{3}{*}{TreeMap}			& 1	& 0.47	& 0.52	& 0.77	& 65		& 70		& 215	& 11		& 11		& 11\\
   		 						& 2	& 8.99	& 9.73	& 4.72	& 1009	& 942	& 3219	& 127	& 122	& 73\\
   								& 3	& -		& -		& 145.56	& -		& -		& 78695	& -		& -		& 887\\

    \hline
  \end{tabular}}
  \caption{Comparing the efficiency of the symbolic heap with the~\gsetxt{} and Lazier\# algorithms.}
  \label{tab:results}
\end{table*}

@@

%\subsection{Analysis}
Although the correctness of~\symtxt{} has been demonstrated,
it remains an open question whether it represents a practical alternative to the
case-splitting methods currently in use. It is possible 
that the benefit of decreasing the number of paths explored is
offset by the increased complexity of the resulting path condition.

To resolve this question, a prototype of the \symtxt{} 
algorithm was developed in the Symbolic PathFinder (SPF)~\cite{DBLP:journals/ase/PasareanuVBGMR13} framework
to evaluate its efficacy on Java data structure artifacts.
%SPF includes an implementation of~\gsetxt{} with lazy initialization.
%In recent work, we implemented the  Lazier and Lazier\#
%algorithms in SPF~\cite{Hillery:2014}.
%
%The prototype of our~\symtxt{} algorithm is implemented as an
%extension to SPF.  
In addition to the operations presented in
this paper, the implementation contains support for operations over
integers, calculating per-path preconditions 
and postconditions, as well as path-coverage-optimal sets of 
test input heaps. 
%A standalone version of the system, independent
%from JPF is also under development~\cite{heapsolver2015}.

The data structures used in the benchmarks are a standard set commonly used in analyses involving heap-manipulating programs~\cite{Deng:2006,Deng:2007,boyapati2002korat,Ferrara:2014,Rosner:2015}, including a linked list, binary search tree, and red/black tree. The BinarySearchTree test uses the class of the same name from Alan Weiss' book~\cite{Weiss:1998}. The test method first asserts a repOk() invariant, before conducting a traversal to find an arbitrary integer key in the tree. The TreeMap method comes from the standard java.util library, and is an implementation of a balanced red/black tree. This method asserts a repOk() invariant, before traversing the tree. The LinkedList test starts by asserting a repOk() invariant, which checks for loops in the linked list using Floyd’s cycle-finding algorithm. The linked list is then traversed using a contains() method that attempts to locate an arbitrary reference key in the list. For benchmarking purposes, the length of reference chain initializations was limited to some integer $k$. For both the tree examples, the number of possible non-isomorphic tree structures grows exponentially in k~\cite{Deng:2007}, resulting a corresponding exponential growth in analysis times for case-splitting techniques.

The results of the experiments are presented in Table~\ref{tab:results}. 
Each row reports the results for the specified $k$-bound 
for each artifact evaluated. The columns show the total
time in seconds, states explored, and paths generated for
each algorithm. The headings \gsetxt{}, L\#, and SH correspond
to the Generalized Symbolic Execution, Lazier\#, and \symtxt{}
algorithms, respectively. The results for
the Lazier algorithm are omitted because they are nearly identical to the Lazier\#
results for these artifacts. A table entry of  '-' indicates the analysis
exceeded the allotted time bound of 1 hour.

Performance improvement ranges from $4.8x$ for
BinarySearchTree at $k$=3, to $118x$ for LinkedList at $k$=6. In fact,
for some $k$-bounds a number of experiments complete exploration
using the \symtxt{} algorithm in a few seconds whereas \gsetxt{} or Lazier\# are
unable to finish, e.g., BinarySearchTree for $k$=4 and LinkedList for
$k$=8. The results demonstrate that for these common data structures,
the complexity of the constraints generated by~\symtxt{} is more 
than offset by the reduction in paths explored.

%how does the per-path cost look for our method

%The number of states explored varies between algorithms. For example,
%\gsetxt{} creates additional points of nondeterminism during field reads, 
%while reference compares are completely deterministic. Thus, in
%example programs with large numbers of reference compares, e.g.,
%TreeMap, state counts for Lazier\# and the \symtxt{} algorithm may
%exceed those for \gsetxt{}. We observed that the additional states
%generated by the symbolic heap algorithm are unsatisfiable at the point
%of reference compares; this is why they do not contribute any
%additional branches in the final symbolic execution tree. Furthermore,
%the since larger state count is essentially a result of bookkeeping, it
%does not impact the overall runtime of the
%technique.

% Some of this discrepancy may be accounted for by the fact that
%\symtxt{} is equipped for incremental solving, however, during
%informal testing with the incremental solver turned off, \symtxt{}
%usually performed better than \gsetxt{} and Lazier\#, especially in
%test runs with high k-bounds.

%The path counts shown in Table~\ref{tab:results} also show the advantage
%of the \symtxt{} algorithm over the \gsetxt{} and L\# algorithms, each of
%which performs case-splitting during the analysis.
%The number of paths explored by the \symtxt{} algorithm is strictly 
%less than or equal to the number
%of paths explored by \gsetxt{} for all artifacts. As a result,
%fewer test cases will be generated to achieve the same control-flow 
%path coverage when compared to~\gsetxt{} and Lazier\#.

%In summary, the benefits of avoiding
%initialization-based non-determinism outweigh the increased
%complexity in the constraints due to the advances made in SMT
%solvers in these examples. In fact, the \symtxt{} algorithm can analyze certain types of programs with
%orders of magnitude greater efficiency than with \gsetxt{} or
%Lazier\#, while covering exactly the same control flow paths.

%We can use this
%result to do more efficient test case generation. From a path
%generated we can use the solutions provided by the constraint solver
%to instantiate a set of concrete heaps to provide as test input.

%Interestingly, \symtxt{} path growth appears
%to be sub-exponential in $k$ for the LinkedList program.

%The additional states generated by summary heap are unsatisfiable
%which is why they do not add nodes in the final symbolic execution
%tree. Furthermore, the larger state count does not impact the overall
%runtime of the technique.

% Despite this, in practice this is often mitigated by the
%fact that \symtxt{} can dispatch address compares without consulting
%the constraint solver.

>>>>

\section{Evaluation}
\label{sec:evaluation}
%\begin{figure*}[t]
\begin{table*} [t]
  \centering
  \caption{Comparing \symtxt{} with the~\gsetxt{} and Lazier\# algorithms.}
  \scalebox{0.88}{\begin{tabular}{| c | c | r | r | r | r | r | r | r | r | r |}
  \hline
   \multirow{2}{*}{Method }&\multirow{2}{*}{ $k$ }
   &\multicolumn{3}{c|}{Time (seconds)} &\multicolumn{3}{c|}{States} &\multicolumn{3}{c|}{ Paths }\\
								&	&\gsetxt{} & L\#		&SL		&\gsetxt{}	& L\# & SL&\gsetxt{}	& L\# 	& SL\\
   \hline
    \multirow{3}{*}{LinkedList }			&3	& 0.91	& 1.21	& 0.69	& 2465	& 2844	& 99		&1656	& 1269	& 25\\
   		 						& 4	& 2.92	& 3.35	& 0.91	& 25774	& 29977	& 155	&17485	& 13550	& 39\\
   								& 5	& 20.78	& 19.47	& 1.59	& 341164	& 400296	& 223	&232743	& 181849	& 56\\
								& 6	& 280.56	& 299.19	& 2.36	&5447980	&6437201	& 303	&3731094	&2933027	& 76\\
								& 7	& -		& -		& 5.07	& -		&-		& 395	&-		&-		& 99\\
								& 8	& -		& -		& 17.49	& -		&-		& 499	&-		&-		& 125\\
								& 9	& -		& -		& 63.96	& -		&-		& 615	&-		&-		& 154\\
								& 10	& -		& -		& 206.93	& -		&-		& 743	&-		&-		& 186\\
    \hline
    \multirow{3}{*}{BinarySearchTree }	& 1	& 0.26	& 0.28	& 0.36	& 19		& 23		& 29		& 6		& 6		& 6\\
   		 						& 2	& 0.83	& 1.28	& 0.93	& 143	& 143	& 145	& 43		& 42		& 33\\
   								& 3	& 20.63	& 25.55	& 4.03	& 1953	& 1703	& 1485	& 515	& 515	& 328\\
								& 4	& -		& -		& 410.89	& -		&-		& 73635	&-		&-		& 15563\\
    \hline
      \multirow{3}{*}{TreeMap}			& 1	& 0.47	& 0.52	& 0.77	& 65		& 70		& 215	& 11		& 11		& 11\\
   		 						& 2	& 8.99	& 9.73	& 4.72	& 1009	& 942	& 3219	& 127	& 122	& 73\\
   								& 3	& -		& -		& 145.56	& -		& -		& 78695	& -		& -		& 887\\

    \hline
  \end{tabular}}

  \label{tab:results}
\end{table*}

@@

%\subsection{Analysis}
The \symtxt{} algorithm is implemented as an extension to the Symbolic
PathFinder (SPF)~\cite{DBLP:journals/ase/PasareanuVBGMR13}
framework. In addition to the operations presented in this paper, the
implementation contains support for operations over integers,
calculating per-path preconditions and postconditions, as well as
generating test input heaps that exercise all feasible control flow
paths. We are currently working on adding support for floating point
operations, arrays, and bit-operations. The \symtxt{} implementation
uses the \texttt{jConstraints}
library~\cite{ase2014-ghilrr,jpf2014-dghirr} to encode the heap
constraints and the z3 solver~\cite{deMouraBjorner08Z3} as the
underlying constraint solver. We leverage the incremental solving
feature in z3 in our experiments. We also implement other
optimizations that cache and re-use solutions to previous
constraints. These optimizations are an important part of our
implementation because only small portions of the heap constraint
change during execution.

SPF includes an implementation of~\gsetxt{} with lazy initialization.
In recent work, we implemented the Lazier and Lazier\# algorithms in
SPF~\cite{Hillery:2014}; these constitute the state of art approaches
case-splitting based lazy initialization techniques. The goal of our
experiment is to evaluate the efficacy of our approach with respect to
other state-of-the-art techniques for symbolic execution of programs
with unbounded complex data input. We empirically evaluate the
\symtxt{} algorithm by considering the following research question:
How does the cost of the \symtxt{} algorithm (SL) compare with that of
\gsetxt{} (GSE), Lazier, Lazier\#(L\#) algorithms?

The \emph{independent} variable in our study is the $k$-bound;
$k$-bounding bounds the length of a reference chain from the root of
the heap~\cite{Kiasan06}.  We select three dependent variables and
measures for our study: (i) time, (ii) states explored, and (iii)
paths generated. The \emph{time} is the total time taken by each
algorithm to explore the symbolic execution tree as well as total
constraint solving time. The \emph{states explored} represents the
number of nodes in the symbolic execution tree, and the \emph{paths
  generated} represents the number of unique paths in the symbolic
execution tree.

The data structures evaluated are a standard set commonly used in
analyses involving heap-manipulating
programs~\cite{Deng:2006,Deng:2007,boyapati2002korat,Ferrara:2014,Rosner:2015},
including a linked list, binary search tree, and red/black tree. We
use a repOk() method (a class invariant) for data structures in
object-oriented code~\cite{boyapati2002korat}. The repOk() methods
provide a mechanism to generate valid inputs for the methods under
test; symbolic execution has been used to generate test inputs using
repOk() methods in~\cite{visser2004test}. In this work, we define and
use repOk() methods to generate valid heaps with the appropriate
$k$-bounds before executing the methods under test. Note that this
allows us to have precision in checking properties of the heap that is
not possible in static analysis based techniques~\cite{Dillig:2011}.

The results of the experiments are presented in
Table~\ref{tab:results}.  Each row reports the results for the
specified $k$-bound for each artifact evaluated. The columns show the
total time in seconds, states explored, and paths generated for each
algorithm. The headings \gsetxt{}, L\#, and SL correspond to the GSE,
Lazier\#, and \symtxt{} algorithms, respectively. The results for the
Lazier algorithm are omitted because they are nearly identical to the
Lazier\# results for these artifacts. A table entry of '-' indicates
the analysis exceeded the allotted time bound of 1 hour.

The number of possible non-isomorphic heap configurations grows
exponentially for case-splitting techniques with a monotonic increase
in $k$, resulting a corresponding exponential increase in analysis
times . This is event in all the examples in
Table~\ref{tab:results}. The GSE and L\# algorithms are unable to
finish exploration in a time bound of one hour for $k\geq 7$ for the
LinkedList, $k\geq 4$ for the BinarySearchTree, and $k\geq 3$ for the
TreeMap examples. The improvement in analysis time for the \symtxt{}
over the state of the art case splitting techniques range from $4.8x$
for BinarySearchTree at $k$=3, to $118x$ for LinkedList at $k$=6. In
fact, for some $k$-bounds a number of experiments complete exploration
using the \symtxt{} algorithm in a few seconds whereas \gsetxt{} or
Lazier\# are unable to finish, e.g., BinarySearchTree for $k$=4 and
LinkedList for $k$=8.

The number of path explored by \symtxt{} algorithm are strictly less
than or equal to the number of paths explored by \gsetxt{} for all the
evaluated artifacts. We can use this result to do more efficient test
case generation. From a generated path and corresponding symbolic heap
we can use the solutions provided by the constraint solver to
instantiate a set of concrete heaps to provide as test input. These
provide a smaller test suite to achieve control-flow path coverage as
compared to GSE and Lazier\#. Note that the ability to perform test
input generation is another advantage of \symtxt{} technique over
static analysis approaches.

The number of states varies between algorithms, for example, \gsetxt{}
has additional points of nondeterminism during field reads, but in
contrast reference compares are completely deterministic. Thus, in
example programs with large numbers of reference compares, such as
TreeMap, state counts for Lazier\# and the \symtxt{} algorithm may
exceed those for \gsetxt{}. We observed that the additional states
generated by the summary heap algorithm are unsatisfiable at the point
of reference compares; this is why they do not contribute any
additional branches in the final symbolic execution tree. Furthermore,
the larger state count and the corresponding satisfiability checks on
the constraint solver does not increase the overall runtime of the
technique.

In summary, the benefits of avoiding case-splitting based
non-determinism outweigh the increased complexity in the constraints
over heaps due to the advances made in SMT solvers in these
examples. In fact, the \symtxt{} algorithm can analyze certain types
of programs with orders of magnitude greater efficiency than with
\gsetxt{} or Lazier\#, while covering exactly the same feasible
control flow paths in the program.

### 2015-09-12 vmcai-2016-tex/intro.tex
Some more changes
edits to evaluation and conclusion

The contribution of this work is a symbolic initialization scheme for uninitialized pointers in a fully symbolic heap. Where GSE lazily instantiates uninitialized pointers to either reference NULL, a new instance of the correct type, or an alias to a previously initialized object, the symbolic initialization in this work creates a guarded value set expressing all of these eventualities in the same symbolic heap. Where GSE branches the search space on each choice, symbolic initialization does not. Symbolic initialization only branches on reference compares, partitioning the symbolic heap to indicate heaps where the references alias and heaps where the references do not.

@@

A proof-of-concept implementation of symbolic initialization is reported in this paper. It demonstrates the efficiency of~\symtxt{}, as compared to existing~\gsetxt{} methods. The use of the guarded value sets does require extra interaction with an SMT solver, but the savings in representing multiple heap shapes in a single representation overcomes the cost of the solver calls. In fact, the comparison shows an exponential increase in the number of heaps that can be analyzed in a set of benchmark examples meaning that other approaches based on~\gsetxt\ fail to complete.

>>>>

The contribution of this work is a symbolic initialization scheme for
uninitialized references in a fully symbolic heap. Where GSE lazily
instantiates uninitialized references to either NULL, a new instance
of the correct type, or an alias to a previously initialized object,
the symbolic initialization in this work creates a guarded value set
expressing all of these eventualities in the same symbolic heap. Where
GSE branches the search space on each choice, symbolic initialization
does not. Symbolic initialization only branches on reference compares,
partitioning the symbolic heap to indicate heaps where the references
alias and heaps where the references do not.

@@

A proof-of-concept implementation of symbolic initialization is
reported in this paper. The implementation is for Java in the Java
Pathfinder tool. It demonstrates the efficiency of~\symtxt{}, as
compared to existing~\gsetxt{} methods. The use of the guarded value
sets does require extra interaction with an SMT solver, but the
savings in representing multiple heap shapes in a single
representation overcomes the cost of the solver calls. In fact, the
comparison shows an exponential increase in the number of heaps that
can be analyzed in a set of benchmark examples meaning that other
approaches based on~\gsetxt\ fail to complete.
